\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles

### 2019-12-26 Response-NAR-Round1.tex

\begin{document}
\maketitle

\tableofcontents

\section{General comments}

First of all, the authors would like to express their deepest gratitude for the quality and thoroughness of the reports written by the reviewers, covering a remarkably comprehensive range of aspects of the submitted manuscript. We have individually addressed each of the \ref{comment:last} comments extracted from the first round of review, each to the best of our ability, but would like to start this response with higher-level answers to recurrent remarks:
\begin{itemize}
	\item It was not within the scope of this research to establish/compare the intrinsic predictive potentials of individual probing reagents/protocols, and we did not make any specific claim in that respect. We understand that some unfortunate presentation choices of  our part may have suggested stronger stances than intended (\emph{e.g.} giving the impression that the performances of SHAPE CE are representative of SHAPE technologies based on RT stops). This was not our intention, or message, and we have attempted in this revision  to correct any such misleading statement, and overall optics;
	\item Nevertheless, all of our analyses, including data produced specifically for the validation of the method, support the notion than combining multiple (3+) probing experiments mitigates the risks of mispredictions due to outliers, and frequently leads to significant improvements as assessed by standard statistical tests suggested by Reviewer \#4;
	\item To the best of our knowledge, \OurTool currently represents the only general method that addresses the automatic modeling of RNA structure from multiple sources of probing data. \OurTool also seems to represent a competitive option for mono-probing predictions, but this observation is merely an element of validation, and (partially) substantiates the hope of an expected good quality of predictions in a multiple setting;
	\item We prefer to process various sources of probing data homogeneously rather than optimize our performances (potentially artificially due to overfitting over scarce data) by learning reagent/protocol specific conversions of reactivities to pseudo-energies. Accordingly, our method only relies on a preliminary calibration with respect to the parameters $m$ and $b$ in the \cite{Deigan2009} pseudo-energy conversion model, which proved robust enough for us not to question this initial choice.
\end{itemize}

Over the course of this extensive round of revisions, we have:
\begin{itemize}
  \item Rewritten extensively our manuscript, notably including the bulk of our method section;
  \item Assessed the significance of our main quantitative claims as per \cite{Xu2012}, and included a new paragraph describing the methodology;
  \item Performed additional tests to ascertain the quality of reactivities produced for \didy: We analyzed the correlations of reactivity profiles at the reactivity level (Supp. Sec. 3.2); We also the sigifnicantly better agreement for all produced reactivity profiles than expected at random (Supp. Sec. 3.1);
  \item Improved the distribution, documentation, and streamlined the installation of \OurTool (although we could not reach a \Software{bioconda} distribution yet);
  \item Fixed numerous typos and unfortunate turns of sentences.
\end{itemize}

We hope that, upon reviews, the editor and reviewers will share our enthusiasm for \OurTool, and consent to its dissemination to the wide readership of the journal.

@@

	{\centering\includegraphics[width=.8\linewidth]{graphs/didy/1M7}\\}

	\end{itemize}

	As can be seen on the right  histograms the read depth is sufficient ($>$10k reads/nucleotides) for these experiments to be interpreted (\emph{ie} for the frequency of mutations to be a good estimate of the probability). The histograms on the left shows that there are significantly more mutations in the modified (denatured and SHAPE) samples than in the untreated ones.

	These histograms are very similar to those deemed as exemplary of a successful experiment by \cite{Smola2015},. 
	In addition, those SHAPE-MaP results are consistent with the X-ray structure (see Supplementary Figure S2 and S3), and their use with our prediction software was shown to improve the accuracy of the output model(s).

@@

	\Comment{Sixth, according to the abstract, IPANEMAP includes an unsupervised learning component. However, if the Deigan’s “m” and “b” parameters were optimized against known structures, then I don’t see which parts are unsupervised. This is another reason why I request to clarify the parameter optimization step. }

	\Answer{
		The distance-based clustering of structures, arguably the most crucial component of \OurTool, is fully unsupervised. We do not believe that a loose calibration of two numerical parameters should raise a reasonable concerns of overfitting.
	}

	\Comment{The bulk of the manuscript describes various benchmarks, typically comparing average accuracies over multiple IPANEMAP runs, over multiple RNAs, or over multiple conditions. However, a comparison of two sets of data points should consider their means in the context of their std’s, otherwise it is often not too informative or conclusive. What I’m missing here are classical tests that quantify the statistical significance associated with each comparison. This is missing in all tests conducted in this work, and I suspect that a few of the reported mean differences might not be very significant. One example of where a test is needed is the comparison to Rsample -  see my comment above. Another example is the comparison to MFE and MEA (Table 2). There are several more. Note that if multiple tests are jointly considered/reported then p-value adjustment (e.g., B-H) might be warranted.}

	\Answer{
	Despite our previously-stated objections, we understand that such metrics are expected for new methods. We assessed the statistical significance of our main assertions using the two-tailed Student $t$-test on paired data with unequal variance, as recommended by \cite{Xu2012}:
	\begin{itemize}
		\item \cite{Hajdin2013} dataset.\\
		Hypothesis $H_0$: \OurTool and \Software{RSample} have similar expected GM\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=6\times 10^{-3}$ $\to$ Rejected at $\alpha=5\times10^{-2}$ 
		\item \cite{Cordero2012} dataset.\\
		  Due to the limited size of the dataset (6 RNAs), none of our observations regarding the average behavior are sufficient to reach statistical significance. We have toned down the title of the corresponding section, and added a remark to the conclusion of the associated results section.
		\item \didy{}  dataset.\\
		Hypothesis $H_0$: \OurTool and MFE have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.2\times 10^{-2}$ $\to$ Borderline rejected at $\alpha=5\times10^{-2}$\\
		Hypothesis $H_0$: \OurTool and MEA have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=14.4\times 10^{-2}$ $\to$ Not significant at $\alpha=5\times10^{-2}$ \\
		Hypothesis $H_0$: Joint analysis of triplet have same MCC as average of three member conditions\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=1.4\times 10^{-9}$ $\to$ Rejected at $\alpha=5\times10^{-2}$ \\
		Hypothesis $H_0$: Joint analysis of triplets and pairs of conditions have same MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.1\times 10^{-2}$ $\to$ Borderline rejected at $\alpha=5\times10^{-2}$ \\
		\item \cite{Cheng2017} dataset.\\
		Hypothesis $H_0$: WT and WT+1M have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=9.7\times10^{-4}$ $\to$ Rejected at $\alpha=5\times10^{-2}$\\ 
		Hypothesis $H_0$: WT+1M and WT+2MC have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.4\times10^{-2}$ $\to$ Rejected at $\alpha=5\times10^{-2}$
	\end{itemize}
	Additional tests can be found in the manuscripts, along with a new section that describes the precise tests carried out and their rationale, further substantiating our claims.
	}

@@

	\Answer{
		It must be noted that we do not base our main conclusion (benefit from the joint analysis of multiple reactivity) solely on the \didy{} case study, but also on the Cordero dataset, where the best performances are also observed upon considering a combination of reactivity profiles associated with DMS, CMCT and NMIA probes. A similar trend is also observed on the \cite{Cheng2017} dataset.
	}

	\Comment{Second, it is my understanding  that no replicate data was generated or analyzed, with the exception of SHAPE-CE. Typically, methods are tested on at least a few RNAs of varying lengths and whenever possible on datasets like Hajdin et al.’s. From such studies, we know that data-directed gains vary substantially between RNAs. For this reason, I would be hesitant to draw any conclusions from probing data for a single RNA.

	That said, I do not expect the authors to generate more data. I do think it should be emphasized that this is merely a case study. The purpose of this study and what one could conclude from it should also be clarified. While the subsection’s title (page 7) indicates this is a case study, I believe that some statements in the Abstract and Discussion require further substantiation and should be toned down.}

	\Answer{We already clarified that this is merely a case study. We wish we could do more (e.g. test on a wider dataset) but given the diversity of probes, a systematic quantitative assessment of individual probes is simply unrealistic, and way beyond the standards enforced for the dissemination of a new {\bf method} in structural bioinformatics.

@@

	However, it must be noted that we do not base our main conclusion (complementarity of probing data, benefit of their integration) solely on the \didy{} case study, but also on the Cordero dataset, where the best performances are also observed upon considering a combination of reactivity profiles associated with DMS, CMCT and NMIA probes. 
	}

	\Comment{The other reason I feel these results should be framed as a case study is that replicate data was not considered in the analysis. The manuscript attempts to differentiate between conditions or between algorithms, but the variation within each condition should be assessed in order to confirm that observed differences are in fact meaningful. }

	\Answer{We understand those reservations but, again, we need to stress that we already  {\bfseries explicitly describe this study as a case study}, and do not make any claim regarding the superiority of a given technology or probe from this study.

	We also did reproduce in triplicate all of our SHAPE CE probing experiments, and observed minor variations at the reactivity level, leaving us to only consider a single representative (average of available reactivities) for each experiment. We now include in this revision the reactivities measured over each of the 3 replicates for each condition.  For SHAPE-MaP protocols,we could not duplicate them due to the relatively high cost and length of those experiments in our wet lab.
	}

@@

	\Comment{The assessment of mono-probing predictions on page 8 does not seem to result in a clear and significant take-home message. Accuracies seem to be comparable between IPANEMAP, MEA, and MFE, but I don’t see what we learn from this. Note also that the data-directed MEA and MFE algorithms were optimized for SHAPE data only, hence they might not do as well on non-SHAPE data (Deng et al., RNA, 2016).}

	\Answer{We agree, and in fact do not draw any conclusion from these results. We only report the MEA/MFE values as references, to support the hypothesis that the MCCs achieved by \OurTool are probably more representative of the reactivity profiles/condition than of a specific predictive model.}

	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols. I don’t know if it was accounted for during normalization step, if normalization was indeed applied.}

	\Answer{We were unsure how to interpret this remark: \Ab{} and \Cb{} are indeed the only reactive nucleotides, so the normalization step is indeed restricted to positions featuring those two nucleotides. If the reviewer is referring to the observations of \cite{Sexton2017} of opposite {\em \Ab{} vs \Cb} biases for RT-based and mutations-based probing respectively, we did not account for it. Indeed, while including corrective terms would seem to represent an opportunity for improved predictions, we would rather treat of reactivity data homogeneously than achieve the best MCCs possible. More details in our response to Comment \ref{arg:homogenous} of Reviewer \#3.
	}

@@

	\Comment{Much space is allocated to outlier detection (cluster B) and then to describing analyses with and without cluster B. I feel that the authors can safely designate cluster B as an outlier and exclude it from the analyses described in the main text. Analyses with cluster B can be described in the Supp.}

	\Answer{Again, we respectfully disagree with this vision. What is important about cluster B is not only the fact that it can be identified from our new metrics (more conclusively than than from the raw reactivities), but also that it can be used to assess the impact of contamination by an erroneous/outlier condition.}

	\Comment{Were all the analyzed datasets normalized by any standard method (e.g., 2\%-8\% or box-plot)? I noticed that the mutate-and-map dataset was renormalized, and to the best of my knowledge, the Hajdin dataset is also normalized. However, I didn’t see any information re: Cordero dataset, and more importantly re: the newly generated data analyzed here. Normalization is especially important in the context of differential analysis between conditions.}

	\Answer{
	All reactivities but the \cite{Cordero2012a} were renormalized using the 2\%-8\% approach of \cite{Deigan2009}.
	For \cite{Cordero2012a}, we used the data directly as we found the reactivities to follow the expected distribution.

	However it is important to note that our analysis does not rely on a direct differential analysis at the reactivity level, except in our newly added Supplementary Section 3.2. The latter features a cross-analysis based on the Pearson correlation, precisely chosen as its results are impervious to affine transforms of the data.
}

### 2019-12-26 IPANEMAP-2019.tex
Updates from Overleaf

	{\centering
			\begin{adjustbox}{max width=\linewidth}
		\begin{tabular}{@{}lccHHH}
			\toprule
			                            &          &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
			Condition                   &  Clust.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
			\OneMSevILUMg               & \Bull{C} &    \Mg    & 85                       & 82      & 82      \\
			\OneMSevILU                 & \Bull{A} &   \NoMg   & 84                       & 82      & 84      \\
			\NMIAMg                     & \Bull{A} &    \Mg    & 81                       & 80      & 80      \\
			\NMIA                       & \Bull{A} &   \NoMg   & 80                       & 69      & 69      \\
			\OneMSevILUThreeMg          & \Bull{C} &    \Mg    & 74                       & 75      & 75      \\
			\NMIAMgCE                   & \Bull{E} &    \Mg    & 73                       & 72      & 73      \\
			\NAIMg                      & \Bull{G} &    \Mg    & 73                       & 68      & 70      \\
			\BzCNMg                     & \Bull{H} &    \Mg    & 71                       & 68      & 54      \\
			\OneMSevMgCE                & \Bull{E} &    \Mg    & 70                       & 69      & 70      \\
			\CMCTMg                     & \Bull{G} &    \Mg    & 70                       & 68      & 73      \\
			\OneMSevILUThree            & \Bull{D} &   \NoMg   & 64                       & 56      & 61      \\
			\DMSMg                      & \Bull{F} &    \Mg    & 63                       & 66      & 67      \\
			\NMIACE                     & \Bull{B} &   \NoMg   & 61                       & 60      & 60      \\
			\OneMSevCE                  & \Bull{B} &   \NoMg   & 60                       & 61      & 61      \\
			\BzCN                       & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\
			\NAICE                      & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\ \midrule
			Avg Technology              &          &  $-$  & 78                       & 74      & 75      \\
			                            &          &  $-$  & 66                       & 65      & 65      \\
			Avg -/+ Mg$^{2+}$           &          &   \NoMg   & 67                       & 64      & 65      \\
			                            &          &    \Mg    & 73                       & 72      & 71.5    \\ \midrule
			{\bfseries Average overall} &          &  $-$  & 70.5                     & 68      & 68.5    \\ \bottomrule
			                            &          &           &                          &
		\end{tabular}
		\end{adjustbox}}

@@

			\multicolumn{4}{@{}l@{}}{Conditions in individual clusters}                                    & \MCB{-NAI} & \MCB{+NAI} \\

@@

	\caption{Predictive performances of conditions across clusters, coupled with impact assessment of NAI.  For each experiment, a single condition is chosen from each cluster except for cluster \Bull{B}. Clusters \Bull{D}, \Bull{F} and \Bull{H}, each consisting of a single condition, contribute their sole condition, leading to the list of conditions  $\mathcal{M}$:=\{\OneMSevILUThree, \BzCNMg and \DMSMg\}. 
		For each combination, the MCC of the predicted structural model is reported, as predicted from the conditions in the presence/absence of \NAICE ('Joint' columns). For each prediction, the average MCC value of mono conditions is reported as a baseline ('Mono' columns).

>>>>

	{\centering
			\begin{adjustbox}{max width=.8\linewidth}
		\begin{tabular}{@{}lccHHH}
			\toprule
			                            &          &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
			Condition                   &  Clust.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
			\OneMSevILUMg               & \Bull{C} &    \Mg    & 85                       & 82      & 82      \\
			\OneMSevILU                 & \Bull{A} &   \NoMg   & 84                       & 82      & 84      \\
			\NMIAMg                     & \Bull{A} &    \Mg    & 81                       & 80      & 80      \\
			\NMIA                       & \Bull{A} &   \NoMg   & 80                       & 69      & 69      \\
			\OneMSevILUThreeMg          & \Bull{C} &    \Mg    & 74                       & 75      & 75      \\
			\NMIAMgCE                   & \Bull{E} &    \Mg    & 73                       & 72      & 73      \\
			\NAIMg                      & \Bull{G} &    \Mg    & 73                       & 68      & 70      \\
			\BzCNMg                     & \Bull{H} &    \Mg    & 71                       & 68      & 54      \\
			\OneMSevMgCE                & \Bull{E} &    \Mg    & 70                       & 69      & 70      \\
			\CMCTMg                     & \Bull{G} &    \Mg    & 70                       & 68      & 73      \\
			\OneMSevILUThree            & \Bull{D} &   \NoMg   & 64                       & 56      & 61      \\
			\DMSMg                      & \Bull{F} &    \Mg    & 63                       & 66      & 67      \\
			\NMIACE                     & \Bull{B} &   \NoMg   & 61                       & 60      & 60      \\
			\OneMSevCE                  & \Bull{B} &   \NoMg   & 60                       & 61      & 61      \\
			\BzCN                       & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\
			\NAICE                      & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\ \midrule
			Avg Technology              &          &  $-$  & 78                       & 74      & 75      \\
			                            &          &  $-$  & 66                       & 65      & 65      \\
			Avg -/+ Mg$^{2+}$           &          &   \NoMg   & 67                       & 64      & 65      \\
			                            &          &    \Mg    & 73                       & 72      & 71.5    \\ \midrule
			{\bfseries Average overall} &          &  $-$  & 70.5                     & 68      & 68.5    \\ \bottomrule
			                            &          &           &                          &
		\end{tabular}
		\end{adjustbox}\\}

@@

			\multicolumn{4}{@{}l@{}}{Conditions in individual clusters ($+\mathcal{M}$)}                                    & \MCB{-NAI} & \MCB{+NAI} \\

@@

	\caption{Predictive performances of conditions across clusters.  In each line, a single condition is chosen from each cluster except for cluster \Bull{B}. \Rev{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average MCC of mono conditions, in all cases in presence/absence of \NAICE.}
